Time invariant analysis of epidemics with \pkg{EpiCompare}

### 2021-05-17 jss/EpiCompare-paper/EpiCompare-paper.tex
section 3 edits - tried again at last two steps in pipeline.

In this section, we present the tools implemented in \pkg{EpiCompare}
and explain how they aid in the data analysis pipeline. In Fig.
\ref{fig:pipeline2}, we
\textcolor{violet}{\sout{illustrate} show}\footnote{\textcolor{violet}{[Ben says: I think illustrate isn't the best adjective here (as it suggests a little more like a guided demonstration, which is a little extreme). I'm happy with other synonyms as well.}}
how our package's functions fit into the data analysis pipeline
introduced in Fig. \ref{fig:pipeline}. All front-facing functions in
\pkg{EpiCompare} are aimed to be as user-friendly as possible. We also
focus on providing the user ``tidyverse'' style functions, that
encourage piping objects from one function to the next and follow clear
``verb'' naming schemes \citep{Wickham2019}. Although users can
incorporate \pkg{EpiCompare} into any step in the data analysis
pipeline, there are two primary points of entry. The first point of
entry is the very beginning with pre-processing and visualizing raw
data, and the second point of entry is after modeling and simulation.
Figure \ref{fig:pipeline2} captures these different paths, and we
highlight how to leverage \pkg{EpiCompare} functionalities in the
subsections below.

@@

The first step of most data analysis is ``cleaning'' the raw data so it
can be explored. Before data can be explored, they must be collected.
Raw data can appear in many different
formats.\footnote{\textcolor{violet}{[Ben says: the previous too sentences are very short and seem choppy when I read them - do they for you as well?]}\textcolor{orange}{I think we can get rid of the 'raw data can appear...' sentence.}}
Sometimes individual records are collected, with times of different
states of the epidemic (infection, recovery, etc.) as well as individual
information like network structure, location, and sub-population
information. Other data collections focus on aggregate counts of
individuals in each epidemic state. In fact, many times only the number
of new infections at each time step (e.g.~weekly case counts) is
observed. Compartment totals (amounts of individuals in each state) are
then imputed from those case counts other information about the disease
and the population of interest. In \pkg{EpiCompare}, we focus on
understanding the overall impact of an outbreak at the
aggregate/population level, which allows for streamlined examination of
overall trends of an epidemic.

To help the practitioner examine epidemics from an aggregate/population
lens, we provide a function called \texttt{agents\_to\_aggregate()}.
This function transforms data about individual/agents' initial entry
into each state (e.g.~start of infection, start of recovery, etc.) to an
aggregate view of how many individuals were in a state at a given time.
There are often situations where
\textcolor{violet}{\sout{aggregating} grouping}\textbackslash footnote\{\textcolor{violet}{[We use aggregating two times in this sentence. I'm open to other synonyms as well, but it seems repetitive. ]}
\textcolor{orange}{It was grouping before ;).  Fine with me.} agents
into subpopulations (e.g.~subpopulations defined by age or sex) can
highlight different aggregate level trends. For example, research by
\citet{rvachev1985,anderson1992,worby2015} develop state-based models
that account for differing disease dynamics in different subpopulations.
In \pkg{EpiCompare}, we facilitate subpopulation analysis by combining
the function \pkg{dplyr}::\texttt{group\_by()} and
\texttt{agent\_to\_aggregate()} to provide aggregation at a
subpopulation
level\footnote{\textcolor{violet}{[Ben says: This sentence seems a bit repetitive with "supopulation". ]}\textcolor{orange}{agreed.  I like 'to provide summaries at a'}}.

Besides aiding subpopulation analysis, \texttt{agents\_to\_aggregate()}
can
\textcolor{violet}{\sout{deal with} accommodate}\footnote{\textcolor{violet}{[Ben says: I think the original word is a bit vague it as "deal with" could mean "accommodate" or "avoid traps associated with", etc. Again, I'm open to other synonyms.]}\textcolor{orange}{I like accomodate}}
a wide range of information about each individual. In fact, this
function can account for infinitely many states. This functionality
allows the practitioner to aggregate information relative to common
states (e.g.~``Susceptible'', ``Infectious'', and ``Recovered'') as well
as more complex states (e.g.~``Exposed'', ``iMmune'', ``Hospitalized'').
Additionally, \texttt{agents\_to\_aggregate()} permits indicators for
death/exit and birth/entry dates. Overall, this function is a powerful
tool for pre-processing data.

\textbf{\textcolor{violet}{Exploratory Data Analysis (}EDA\textcolor{violet}{)}}

In the early stages of a project, familiarizing oneself with the data
usually means figuring out useful combinations of visualizations and
numerical summaries of the data both at population and subpopulation
level. An expert coder can start with \texttt{agents\_to\_aggregate()}
to successfully accomplish
\textcolor{violet}{exploratory data analysis (}EDA\textcolor{violet}{)}
in many ways. \pkg{EpiCompare} also includes tools that allow a novice
coder to rapidly explore data, provided there \textcolor{violet}{are}
three unique epidemiological states (like in the SIR model). Building on
\pkg{ggplot2} and \pkg{ggtern} packages, \pkg{EpiCompare}'s
\texttt{geom\_aggregate()} provides a
\textcolor{violet}{\sout{rapid}}\footnote{\textcolor{violet}{[Ben says: this word is used in the sentence before, and I don't think it adds too much to this sentence.]} \textcolor{orange}{accepted}}
way to explore how different subpopulations experience of an epidemic
\citep{Wickham2016, Hamilton2018}. The function
\texttt{geom\_aggregate()} provides a visualization tool to holistically
examine aggregate level information across different subpopulations by
visualizing each subpopulation's epidemic trajectory in the
three-dimensional state space. Visualization tools for three-state
models were developed because (1) SIR models are some of the most common
and basic epidemic state-based models and (2) our three-dimensional
simplex representation of these epidemics emphasizes a
\textcolor{violet}{\sout{"time-invarance"}time-invariant}\footnote{\textcolor{violet}{[Ben says: I don't think we really need quotation marks here as we've discussed the idea in the previous sections.]} \textcolor{orange}{agreed}}
representation of the data (for a refresher see Section
\ref{sec:time-invariant}).

@@

After getting a sense of what a past or current epidemic looks like with
EDA, \textcolor{violet}{\sout{often}} the next step in the data analysis
pipeline is
\textcolor{violet}{often}\footnote{\textcolor{violet}{[Ben says: By moving this word here I found, as a reader, I didn't have as many pauses while reading this sentence. Maybe it's just me.]}}
model fitting and/or simulation. In this step and the next step
(post-processing), we highlight how practitioners can pair models and
simulations of epidemics from outside of \pkg{EpiCompare} with analysis
and simulation tools in \pkg{EpiCompare}. While \pkg{EpiCompare} does
not focus on fitting models to data, we do provide some flexible
functions for simulation of basic discrete-time epidemic-state models.
These functions simulate individual-level information based on
practitioner estimated transition rates between states and can be
combined with \texttt{agents\_to\_aggregate()} to view these simulations
through an aggregate lens. The function \texttt{simulate\_SIR\_agents()}
simulates a basic SIR epidemic with user inputs for the number of
simulations, the initial number in each state, the infection and
recovery parameters \((\beta, \gamma\)), and the total number of
discrete time steps. This function allows for easy access to SIR model
analysis and comparison. Beyond SIR models, the function
\texttt{simulate\_agents()} takes as input a user-specified
state-transition matrix and other epidemic parameters to allow the user
to create simulations for an outbreak with \textit{any} number of states
and any number of transitions among them. This flexibility in states can
be used to also reflect group-based dynamics. In turn, this allows for
users to explore the space of models in an intuitive way without getting
bogged down by too much mathematical detail. For consistency, we have
made output from \texttt{simulate\_agents()} and
\texttt{simulate\_SIR\_agents()} compatible with
\texttt{agents\_to\_aggregate()} so aggregate information may easily be
accessed.

@@

\footnote{\textcolor{violet}{[Ben says: I found the structure of this section slightly confusing. It appears that the 2 subsections are (1) model based transformation and (2) simulation based transformations. The fact that there are 2 points isn't quickly apparent when reading it, and I found it odd that the 3rd sentence of the first paragraph introduces model processing (which the second paragraph also discusses).]}\textcolor{orange}{agreed that this section needs some work.  but to me those aren't the clear points, it seems like (1) put external model results in the same format and (2) transform model output to filaments.  that said i don't see how we're over-selling the package.  would 'transform and disseminate' be better?}}
If practitioners wish to compare models-to-observations or even
models-to-models, they needs to post-process their models and
simulations to disseminate the results in an easily digestible format.
In general, post-processing of modeling and simulation consists of
making summary statistics, plots, and tables. Model output summaries can
be very complex, and as a result, a number of epidemic modeling
\proglang{R} packages return a special class. The special classes often
contain a plethora of information about residuals, model diagnostics,
input parameters, and more. While incredibly useful, these special
classes can be difficult for novice coders to handle.

@@

@@

Finally, in \pkg{EpiCompare} we provide a set of comparison and
assessment tools for model and simulation results that extend beyond the
standard performance metrics (e.g.~mean squared error or AIC).
Corresponding\footnote{\textcolor{violet}{[Ben says: "corresponding to" doesn't seem to be the correct phrase here. If we wanted to use "corresponding" I might imagine saying "corresponding to the araguments for time-invariant analysis in Section 2.2, EpiCompare provides such tools." I'm open to keeping it if you think it's the perfect word choice.]}\textcolor{orange}{"To support the ideas presented in Sec."?}}
to Section \ref{sec:beyond-r0-sir}, \pkg{EpiCompare} provides a set of
time-invariant tools to compare and evaluate epidemic models and
simulations. We have found these tools to be specifically applicable for
situations where only one ``season'' or
``cycle''\footnote{\textcolor{violet}{[Ben says: I think we can remove the paratheses here.]}}
of the epidemic has occurred.

\footnote{\textcolor{violet}{[Ben says: multiple parts of this paragraph might benefit with capturing a bit more of the motivation in the rewrite of Section 2 for the tools we're presenting.]}\textcolor{orange}{I'm not saying 'no' but I'd do it in an appendix and not the main text.  I don't think equations belong in Section 3.}}One
tool we provide to assess models is through the creation of geometric
prediction regions, which are useful when we treat epidemics like
filaments. If there is a set of simulated epidemics from a model, we can
create a geometric prediction region for the expected trajectory of the
epidemic in the state space. For three-state epidemic models, we provide
the \texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction to the true epidemic
trajectory or comparing the prediction regions defined by two different
models' simulations can be done
\textcolor{violet}{by eye.\sout{with the 'eye test.'}} In
\pkg{EpiCompare} we also provide these prediction regions for epidemic
models with with more than three states. The functions
\texttt{create\_convex\_hull\_structure()} and
\texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}) and the ability to
assess the Hausdorff distance between prediction regions based on
simulations from different model (\texttt{hausdorff\_dist()}).

\footnote{\textcolor{violet}{[Ben says: random thought (not needed) - maybe we do provide some equations for these measures and also maybe mathematically express how we'd examine extremeness?]}}We
also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}. Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations.

@@

\footnote{\textcolor{violet}{[Ben says: Shannon, would you mind reading this whole section over again once we've finished edits for section 2 and 3? This initial paragraph seems to be stating section 3's story.]}}In
this section, we highlight many of the tools available in
\pkg{EpiCompare}. As previously discussed, these tools include data
cleaning; visualization; modeling and simulation; post-processing; and
comparison and model assessment, in accordance with the data analysis
pipeline (Fig. \ref{fig:pipeline}). We show a full data analysis from
beginning to end that can be accomplished in a streamlined and
standardized manner via \pkg{EpiCompare}.

>>>>

In this section, we present the tools implemented in \pkg{EpiCompare}
and explain how they aid in the data analysis pipeline. In Fig.
\ref{fig:pipeline2}, we
show\footnote{\textcolor{violet}{[Ben says: I think illustrate isn't the best adjective here (as it suggests a little more like a guided demonstration, which is a little extreme). I'm happy with other synonyms as well.} \textcolor{orange}{accept}}
how our package's functions fit into the data analysis pipeline
introduced in Fig. \ref{fig:pipeline}. All front-facing functions in
\pkg{EpiCompare} are aimed to be as user-friendly as possible. We also
focus on providing the user ``tidyverse'' style functions, that
encourage piping objects from one function to the next and follow clear
``verb'' naming schemes \citep{Wickham2019}. Although users can
incorporate \pkg{EpiCompare} into any step in the data analysis
pipeline, there are two primary points of entry. The first point of
entry is the very beginning with pre-processing and visualizing raw
data, and the second point of entry is after modeling and simulation.
Figure \ref{fig:pipeline2} captures these different paths, and we
highlight how to leverage \pkg{EpiCompare} functionalities in the
subsections below.

@@

The first step of most data analysis is ``cleaning'' the raw data so it
can be explored. Before data can be explored, they must be collected.
\sout{Raw data can appear in many different formats.}\footnote{\textcolor{violet}{[Ben says: the previous too sentences are very short and seem choppy when I read them - do they for you as well?]}\textcolor{orange}{I think we can get rid of the 'raw data can appear...' sentence.} RESOLVED 5/17}
Sometimes individual records are collected, with times of different
states of the epidemic (infection, recovery, etc.) as well as individual
information like network structure, location, and sub-population
information. Other data collections focus on aggregate counts of
individuals in each epidemic state. In fact, many times only the number
of new infections at each time step (e.g.~weekly case counts) is
observed. Compartment totals (amounts of individuals in each state) are
then imputed from those case counts other information about the disease
and the population of interest. In \pkg{EpiCompare}, we focus on
understanding the overall impact of an outbreak at the
aggregate/population level, which allows for streamlined examination of
overall trends of an epidemic.

To help the practitioner examine epidemics from an aggregate/population
lens, we provide a function called \texttt{agents\_to\_aggregate()}.
This function transforms data about individual/agents' initial entry
into each state (e.g.~start of infection, start of recovery, etc.) to an
aggregate view of how many individuals were in a state at a given time.
There are often situations where
\textcolor{violet}{\sout{aggregating} grouping}\footnote{RESOLVED 5/17 (aggregating)}
agents into subpopulations (e.g.~subpopulations defined by age or sex)
can highlight different aggregate level trends. For example, research by
\citet{rvachev1985,anderson1992,worby2015} develop state-based models
that account for differing disease dynamics in different subpopulations.
In \pkg{EpiCompare}, we facilitate subpopulation analysis by extending
\texttt{agents\_to\_aggregate()} to work with
\pkg{dplyr}::\texttt{group\_by()}, the tidyverse approach to the
split-apply-combine paradigm for data analysis
\citep{wickham2011}.\textbackslash footnote\{Changed to try to clarify
how how agents\_to\_aggregate works with group\_by and why
(split-apply-combine)\}

\textcolor{orange}{old text below}

{[}by combining the function \pkg{dplyr}::\texttt{group\_by()} and
\texttt{agents\_to\_aggregate()} to provide aggregation at a
subpopulation
level\footnote{\textcolor{violet}{[Ben says: This sentence seems a bit repetitive with "supopulation". ]}\textcolor{orange}{agreed. aggregation at a group level}}.{]}

Besides aiding subpopulation analysis, \texttt{agents\_to\_aggregate()}
can
accommodate\footnote{\textcolor{violet}{[Ben says: I think the original word is a bit vague it as "deal with" could mean "accommodate" or "avoid traps associated with", etc. Again, I'm open to other synonyms.]}\textcolor{orange}{I like accomodate}  RESOLVED 5/17}
a wide range of information about each individual. In fact, this
function can account for infinitely many states. This functionality
allows the practitioner to aggregate information relative to common
states (e.g.~``Susceptible'', ``Infectious'', and ``Recovered'') as well
as more complex states (e.g.~``Exposed'', ``iMmune'', ``Hospitalized'').
Additionally, \texttt{agents\_to\_aggregate()} permits indicators for
death/exit and birth/entry dates. Overall, this function is a powerful
tool for pre-processing data.

**Exploratory Data Analysis (EDA)*

In the early stages of a project, familiarizing oneself with the data
usually means figuring out useful combinations of visualizations and
numerical summaries of the data both at population and subpopulation
level. An expert coder can start with \texttt{agents\_to\_aggregate()}
to successfully accomplish exploratory data analysis (EDA) in many ways.
\pkg{EpiCompare} also includes tools that allow a novice coder to
rapidly explore data, provided there \textcolor{violet}{are} three
unique epidemiological states (like in the SIR model). Building on
\pkg{ggplot2} and \pkg{ggtern} packages, \pkg{EpiCompare}'s
\texttt{geom\_aggregate()} provides a
\footnote{\textcolor{violet}{[Ben says: this word is used in the sentence before, and I don't think it adds too much to this sentence.]} \textcolor{orange}{accepted}. RESOLVED 5/17}
way to explore how different subpopulations experience of an epidemic
\citep{Wickham2016, Hamilton2018}. The function
\texttt{geom\_aggregate()} provides a visualization tool to holistically
examine aggregate level information across different subpopulations by
visualizing each subpopulation's epidemic trajectory in the
three-dimensional state space. Visualization tools for three-state
models were developed because (1) SIR models are some of the most common
and basic epidemic state-based models and (2) our three-dimensional
simplex representation of these epidemics emphasizes a
time-invariant\footnote{\textcolor{violet}{[Ben says: I don't think we really need quotation marks here as we've discussed the idea in the previous sections.]} \textcolor{orange}{agreed} RESOLVED 5/17}
representation of the data (for a refresher see Section
\ref{sec:time-invariant}).

@@

After getting a sense of what a past or current epidemic looks like with
EDA, the next step in the data analysis pipeline is
often\footnote{\textcolor{violet}{[Ben says: By moving this word here I found, as a reader, I didn't have as many pauses while reading this sentence. Maybe it's just me.]}RESOLVED 5/17}
model fitting and/or simulation. In this step and the next step
(post-processing), we highlight how practitioners can pair models and
simulations of epidemics from outside of \pkg{EpiCompare} with analysis
and simulation tools in \pkg{EpiCompare}. While \pkg{EpiCompare} does
not focus on fitting models to data, we do provide some flexible
functions for simulation of basic discrete-time epidemic-state models.
These functions simulate individual-level information based on
practitioner estimated transition rates between states and can be
combined with \texttt{agents\_to\_aggregate()} to view these simulations
through an aggregate lens. The function \texttt{simulate\_SIR\_agents()}
simulates a basic SIR epidemic with user inputs for the number of
simulations, the initial number in each state, the infection and
recovery parameters \((\beta, \gamma\)), and the total number of
discrete time steps. This function allows for easy access to SIR model
analysis and comparison. Beyond SIR models, the function
\texttt{simulate\_agents()} takes as input a user-specified
state-transition matrix and other epidemic parameters to allow the user
to create simulations for an outbreak with \textit{any} number of states
and any number of transitions among them. This flexibility in states can
be used to also reflect group-based dynamics. In turn, this allows for
users to explore the space of models in an intuitive way without getting
bogged down by too much mathematical detail. For consistency, we have
made output from \texttt{simulate\_agents()} and
\texttt{simulate\_SIR\_agents()} compatible with
\texttt{agents\_to\_aggregate()} so aggregate information may easily be
accessed.

@@

If practitioners wish to compare models-to-observations or even
models-to-models, they need to post-process their models and simulations
to disseminate the results in an easily digestible format. In general,
post-processing of modeling and simulation consists of making summary
statistics, plots, and tables. In \pkg{EpiCompare}, we provide (1)
functions to standardize simulation and model output from external
packages and (2) a function to transform standardized simulation and
model output into a format amenable to time-invariant
analysis.\footnote{new signpost.  There are two paragraphs below corresponding to the two points.  The second paragraph could use another sentence maybe.  I tried to emphasize the process of output -> standard format -> transformation to time invariant analysis.}

Modeling and simulation output can be very complex objects, and as a
result, a number of epidemic modeling \proglang{R} packages return a
special class. The special classes often contain a plethora of
information about residuals, model diagnostics, input parameters, and
more. While incredibly useful, these special classes can be difficult
for novice coders to handle.
\textcolor{orange}{motivates for fortify but not filamental compression}.
To this end, \pkg{EpiCompare} provides a series of
\texttt{fortify}-style methods, called \texttt{fortify\_aggregate()}
which transform output from infectious disease modeling and simulation
packages like \pkg{pomp} and \pkg{EpiModel} into tidy-styled data frames
which contain information about the total number of individuals in each
state at a given time, for a given simulation. These fortify functions
have output that is consistent with that of
\texttt{agents\_to\_aggregate()}. These standardized outputs can then be
piped to summaries, tables, and plots.

Because the fortified output contains temporal information and we
provide the function, \texttt{filament\_compression()}, to transform
temporally defined epidemics to their filamental representations. These
filaments can then be fairly compared to one another or passed to
further time-invariant analysis tools described below.

{[}{[}\textcolor{orange}{OLD TEXT BELOW}

\footnote{\textcolor{violet}{[Ben says: I found the structure of this section slightly confusing. It appears that the 2 subsections are (1) model based transformation and (2) simulation based transformations. The fact that there are 2 points isn't quickly apparent when reading it, and I found it odd that the 3rd sentence of the first paragraph introduces model processing (which the second paragraph also discusses).]}\textcolor{orange}{agreed that this section needs some work.  but to me those aren't the clear points, it seems like (1) put external model results in the same format and (2) transform model output to filaments.  that said i don't see how we're over-selling the package.  would 'transform and disseminate' be better?}}
If practitioners wish to compare models-to-observations or even
models-to-models, they needs to post-process their models and
simulations to disseminate the results in an easily digestible format.
In general, post-processing of modeling and simulation consists of
making summary statistics, plots, and tables.

Model output summaries can be very complex, and as a result, a number of
epidemic modeling \proglang{R} packages return a special class. The
special classes often contain a plethora of information about residuals,
model diagnostics, input parameters, and more. While incredibly useful,
these special classes can be difficult for novice coders to handle.
\textcolor{orange}{motivates for fortify but not filamental compression}

@@

{]}{]}

@@

Finally, the data analysis pipeline often ends with plots, tables, and
summary statistics that are used to compare different models, and
simulations to one another. In \pkg{EpiCompare} we provide a set of
comparison and assessment tools for model and simulation results that
extend beyond the standard performance metrics (e.g.~mean squared error
or AIC) and into the lens of time-invariant
analysis.\footnote{a bit of an awk transition but I wanted to use past wording about TIA.}
We have found these tools to be specifically applicable for situations
where only one ``season'' or ``cycle'' of an epidemic has occurred.

One tool we provide to assess models is through the creation of
geometric prediction regions, which are useful when we treat epidemics
like filaments. If there is a set of simulated epidemics from a model,
we can create a geometric prediction region for the expected trajectory
of the epidemic in the state space. For three-state epidemic models, we
provide the \texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction to the true epidemic
trajectory or comparing the prediction regions defined by two different
models' simulations can be done by eye. In \pkg{EpiCompare}, we also
provide these prediction regions for epidemic models with with more than
three states. The functions \texttt{create\_convex\_hull\_structure()}
and \texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}) and the ability to
assess the Hausdorff distance between prediction regions based on
simulations from different model (\texttt{hausdorff\_dist()}). The
vignettes for \texttt{EpiCompare} go into much more mathematical detail
about how these functions can be used in conjunction with time-invariant
analysis.

{[}{[}\textcolor{orange}{OLD TEXT}

Finally, in \pkg{EpiCompare} we provide a set of comparison and
assessment tools for model and simulation results that extend beyond the
standard performance metrics (e.g.~mean squared error or AIC).
Corresponding\footnote{\textcolor{violet}{[Ben says: "corresponding to" doesn't seem to be the correct phrase here. If we wanted to use "corresponding" I might imagine saying "corresponding to the araguments for time-invariant analysis in Section 2.2, EpiCompare provides such tools." I'm open to keeping it if you think it's the perfect word choice.]}\textcolor{orange}{"To support the ideas presented in Sec."?}}
to Section \ref{sec:beyond-r0-sir}, \pkg{EpiCompare} provides a set of
time-invariant tools to compare and evaluate epidemic models and
simulations. We have found these tools to be specifically applicable for
situations where only one ``season'' or
``cycle''\footnote{\textcolor{violet}{[Ben says: I think we can remove the paratheses here.]} \textcolor{orange}{accepted}}
of the epidemic has occurred.

\footnote{\textcolor{violet}{[Ben says: multiple parts of this paragraph might benefit with capturing a bit more of the motivation in the rewrite of Section 2 for the tools we're presenting.]}\textcolor{orange}{motivate better}}One
tool we provide to assess models is through the creation of geometric
prediction regions, which are useful when we treat epidemics like
filaments. If there is a set of simulated epidemics from a model, we can
create a geometric prediction region for the expected trajectory of the
epidemic in the state space. For three-state epidemic models, we provide
the \texttt{ggplot}/\texttt{ggtern} extension
\texttt{geom\_prediction\_band()} which creates a prediction region
around the top \(1-\alpha\) proportion of the simulations. In this
visual setting, comparing this prediction to the true epidemic
trajectory or comparing the prediction regions defined by two different
models' simulations can be done
\textcolor{violet}{by eye.\sout{with the 'eye test.'}\textcolor{orange}{I'm not saying 'no' but I'd do it in an appendix and not the main text.  I don't think equations belong in Section 3.}}
In \pkg{EpiCompare} we also provide these prediction regions for
epidemic models with with more than three states. The functions
\texttt{create\_convex\_hull\_structure()} and
\texttt{create\_delta\_ball\_structure()} create different geometric
representations of prediction regions for any dimensional state-based
model. For both of these geometric structures, we provide functions to
check if a path is contained (\texttt{contained()}) and the ability to
assess the Hausdorff distance between prediction regions based on
simulations from different model (\texttt{hausdorff\_dist()}).

\footnote{\textcolor{violet}{[Ben says: random thought (not needed) - maybe we do provide some equations for these measures and also maybe mathematically express how we'd examine extremeness?]}}We
also provide functions to calculate the ``extremeness'' of a true
epidemic trajectory compared to simulated epidemics via the
equi-distance filamental trajectory representation as mentioned in
Section \ref{sec:beyond-r0-sir}. Specifically, functions like
\texttt{distance\_pseudo\_density\_function()} can calculate a
pseudo-density estimate of the true epidemic relative to simulated ones.
Functions \texttt{distance\_depth\_function()} and
\texttt{local\_distance\_depth\_function()} provide depth scores that
suggest how geometrically central an epidemic is to simulations. {]}{]}

@@

Finally, in this section we show how tools from \pkg{EpiCompare} can be
used in each step of the data analysis pipeline shown in Fig.
\ref{fig:pipeline}. We analyze an outbreak of measles in the town of
Hagelloch, Germany from 1861-1862, a data set organized by
\cite{pfeilsticker1863}. The data was later made visible by
\cite{oesterle1992} and made available in an \proglang{R} by
\cite{surveillance2017}. The Hagelloch data includes a rich set of
features including household members, school level, household locations,
date of first symptoms (prodromes), date of measles rash, and even the
alleged infector. A subset of the data is shown in Table
\ref{tab:hags-people}. Because of these rich features, this data set has
been an ideal testing ground methodology in infectious disease
epidemiology and is used in work by
\cite{Neal2004,britton2011,groendyke2012,becker2016}.

{[}{[}OLD TEXT

\footnote{\textcolor{violet}{[Ben says: Shannon, would you mind reading this whole section over again once we've finished edits for section 2 and 3? This initial paragraph seems to be stating section 3's story.]}}In
this section, we highlight many of the tools available in
\pkg{EpiCompare}. As previously discussed, these tools include data
cleaning; visualization; modeling and simulation; post-processing; and
comparison and model assessment, in accordance with the data analysis
pipeline (Fig. \ref{fig:pipeline}). We show a full data analysis from
beginning to end that can be accomplished in a streamlined and
standardized manner via \pkg{EpiCompare}. {]}{]}
